---
title: "ds202_lab4"
author: "Maxwell Farver"
date: "March 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(tidyverse)
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
str(defense)
```

```{r}
offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
str(offense)
```


```{r}
biography <- read_excel('cyclonesFootball2018.xlsx', sheet='Biography')
str(biography)
```

###Cleaning###

1)
```{r}
player_names <- factor(biography$Name)
opponent_names <- factor(unique(defense$Opponent_Opponent))
```

2)
```{r}
offClean <- offense %>% separate(`Passing_CMP-ATT`, into = c("cmp", "att"), sep = '-')

offClean$cmp <- str_replace(offClean$cmp, ' ', '')

offClean$att <- str_replace(offClean$att, ' ', '')

offClean <- offClean %>% mutate(Receiving_REC = as.numeric(Receiving_REC))

offClean <- offClean %>% mutate(Receiving_YDS = as.numeric(Receiving_YDS))

offClean <- offClean %>% mutate(Receiving_TD = as.numeric(Receiving_TD))

offClean <- offClean %>% mutate(Rushing_ATT = as.numeric(Rushing_ATT))

offClean <- offClean %>% mutate(Rushing_YDS = as.numeric(Rushing_YDS))

offClean <- offClean %>% mutate(Rushing_TD = as.numeric(Rushing_TD))

offClean <- offClean %>% mutate(cmp = as.numeric(cmp))

offClean <- offClean %>% mutate(att = as.numeric(att))

offClean <- offClean %>% mutate(Passing_YDS = as.numeric(Passing_YDS))

offClean <- offClean %>% mutate(Passing_TD = as.numeric(Passing_TD))

offClean <- offClean %>% mutate(Passing_INT = as.numeric(Passing_INT))
```

```{r}
defClean <- defense
defClean <- defClean %>% mutate(Tackles_Solo, Tackles_Solo = as.numeric(Tackles_Solo))
defClean <- defClean %>% mutate(Tackles_ASST, Tackles_ASST = as.numeric(Tackles_ASST))
defClean <- defClean %>% mutate(Tackles_TFL, Tackles_TFL = as.numeric(Tackles_TFL))
defClean <- defClean %>% mutate(Tackles_Sack, Tackles_Sack = as.numeric(Tackles_Sack))
defClean <- defClean %>% mutate(Turnover_FF, Turnover_FF = as.numeric(Turnover_FF))
defClean <- defClean %>% mutate(Turnover_FR, Turnover_FR = as.numeric(Turnover_FR))
defClean <- defClean %>% mutate(Turnover_INT, Turnover_INT = as.numeric(Turnover_INT))
defClean <- defClean %>% mutate(Pass_QBH, Pass_QBH = as.numeric(Pass_QBH))
defClean <- defClean %>% mutate(Pass_PB, Pass_PB = as.numeric(Pass_PB))
```

3)
```{r}
biography <- biography %>% separate(Height, into = c('ft', 'inch'), sep = '-') %>%
  mutate(Height = as.numeric(ft)*12 + as.numeric(inch))

biography <- subset(biography, select = -c(ft, inch))
```

```{r}
offClean
defClean
biography
```

###Tidying###

1)
```{r}
defClean <- defClean %>% gather(key=Statistic, value=Recorded, 3:length(defClean))
defClean
```

2)
```{r}
defClean %>% 
  ggplot(aes(x=Recorded)) + geom_histogram() +
  facet_wrap(~Statistic, scales="free") + geom_abline(colour = "grey50")
```

Turnover_FR, Pass_PB, Pass_QBH, Turnover_INT. Turnover_FF, Tackles_Sack

3)
```{r}
WV <- defClean %>%
  filter(Opponent_Opponent == 'West Virginia', Statistic == 'Tackles_Solo')

KS <- defClean %>%
  filter(Opponent_Opponent == 'Kansas State', Statistic == 'Tackles_Solo')

tackles <- inner_join(WV, KS, by='Name')
str(tackles)

ggplot(tackles, aes(x=`Recorded.x`, y = `Recorded.y`)) + geom_point() +
  xlab('West Virignia') +
  ylab('Kansas State') +
  ggtitle('Comparison between K State and West Virginia Games')
```

We played better against Kansas State, there were fewer tackles made on average against West Virginia per player this season.

####Part 2####
1)
```{r}
biography <- biography %>% separate(Hometown, into= c('City', 'State'), sep=', ')
```

2)
```{r}
table(biography$State)
```

####Part 3####

1)
```{r}
joined <- inner_join(biography, offClean, by = "Name")
str(joined)
```

2)
```{r}
ggplot(joined, aes(x=Weight, y=Receiving_YDS)) + geom_point()
```

There seems to be a trend of lower weight athletes having more receiving yard. This is disregarding 225 as it leads me to believe it to be a singular player with multiple high-yardage games.

3)
```{r}
outlier <- biography %>%
  filter(Weight == 225)

outlier
```

I was curious to seee whether or not my guess was corrrect. I checked to see who weighed 225 during the season and sure enough Hakeem Butler was one of the two players. Butler is in the top 5 all time for receiving yards in Iowa State history, so he skews the graph fairly well.